Complete Real-Time Liver Model including Glisson Capsule, Vascular Network and Parenchyma

### 2013-02-28 Doc/Miccai2013/miccai2013.tex
paper
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@524 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\begin{document}
%
%\frontmatter          % for the preliminaries
%
\mainmatter              % start of the contributions
%
\title{Complete Liver Model for Real-time Simulation} % TODO
% Or: Modelling the Glisson's Gapsule in Real-time Simulation of Liver
% or anything else ...
%
\titlerunning{TODO: Title}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Anonymous}
%\author{%
%Tom\'a\v{s} Golembiovsk\'y\inst{1,2} \and%
%Igor Peterl\'ik\inst{3} \and%
%Christian Duriez\inst{2} \and%
%Stephan\'e Cotin\inst{2,3}%
%}
%
%\authorrunning{Tom\'a\vs Golembiovsk\'y et al.} % abbreviated author list (for running head)
%
%\institute{%
%Faculty of Informatics, Masaryk University, Brno, Czech Republic\and%
%INRIA Lille -- Nord, France\and%
%IHU, Strasbourg, France%
%}
%
\maketitle

@@

\keywords{liver, Glisson's capsule, deformation modeling, membrane}
\end{abstract}

@@

\CD{This paragraph may change if we use an other model... but maybe we should directly say that we rely on vascularized model ?}
It is known that the parenchyma exhibits viscoelastic behaviour \TG{Cite
something}, we however employ simpler elastic model as we are not so much
interested in time-dependent behaviour but rather in static equilibrium
under certain conditions.
%
The parenchyma is modeled using corotational finite elements. The corotational
formulation allows us to deal with geometrical non-linearities but still
rely on linear stress-strain relationship. Thus rotations and large
deformations are possible. In the corotational formulation the deformation
of each element is expressed in the local frame of reference. There are
several ways of computing the corotational frame for elements, we rely on
the geometrical method proposed in \cite{Nesme2005}.
% NOTE: This description of corotational method is very simplified and could be extended.
The stiffness matrix $\Mat{K}$ can be expressed in terms of the deformation
$u$ and the force as $\Vec{f} = \Mat{K}(\Vec{u}) \Vec{u}$. Where the
components of $\Mat{K}$ depend on the orientation of each element in the
current simulation step.

@@

\subsection{Blood Vessels} %{{{

\CD{Just a proposal: put this subsection before Glisson Capsule}
We also combined our composite model with vascularization method proposed
by Peterl\'{i}k et al. \cite{Peterlik2012}. The method is based on linear beams
with local frames of reference (in a many aspects similar to the
corotational formulation). As such the model also handles geometric
non-linearities in the deformation. The proposed model assumes no relative
motion of vessels to the parenchyma. Segmented vascular network is
discretized into the set of linear elements that are then projected into
the tetrahedral mesh of the parenchyma. During the simulation forces and
torques resulting from the deformations of vessels by parenchyma are
computed and propagated back onto the tetrahedral mesh.

@@

\CD{If the experiments demonstrate that the stiffness of the material is different, the thickness is very very small... so we have to evaluate if it is really necessary to include the model of the capsule for both local and global deformations.}

\subsection{Local deformations} %{{{

\CD{Maybe explain why we call that "local deformations" and in which case it is important to capture such local deformations (i.e. insertion of needle, contacts with a sharp instrument...).}

Hollenstein et al. \cite{Hollenstein2006} measured the response of the
bovine liver to the negative pressure using their aspiration test. One of
the observations was that the Glissons capsule has non-negligible influence
and modeling only the parenchyma leads to overestimation of the
deformation. We reproduced the test with our model and acquired similar
results.

The aspiration device consists of a tube with 1 cm in diameter and is able
to control internal pressure in the tube. The test is performed by
attaching the tube to the tissue and meassuring the tissue response. We
have set up a simulation in the spirit of this test. 
% TODO: fix the scene! tube has 5 mm radius, not 1 cm
The simulation is performed with a 3x3x3 cm cube of liver tissue and tube
with 1 cm radius. The pressure of 2 kPa is simulated inside the tube.

Figure \ref{fig-aspiration} shows the profiles of cuts in the middle of the
cube for the model with and without capsule. It can be seen that the
deformation for the model without capsule is much larger which is in
aggreement with the data obtained by Hollenstein et al.

>>>>

\begin{document}
%
%\frontmatter          % for the preliminaries
%
\mainmatter              % start of the contributions
%
\title{Complete Real-Time Liver Model including Glisson Capsule, Vascular Network and Parenchyma}
%
\titlerunning{TODO: Title}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Anonymous}
%\author{%
%Tom\'a\v{s} Golembiovsk\'y\inst{1,2} \and%
%Igor Peterl\'ik\inst{3} \and%
%Christian Duriez\inst{2} \and%
%Stephan\'e Cotin\inst{2,3}%
%}
%
%\authorrunning{Tom\'a\vs Golembiovsk\'y et al.} % abbreviated author list (for running head)
%
%\institute{%
%Faculty of Informatics, Masaryk University, Brno, Czech Republic\and%
%INRIA Lille -- Nord, France\and%
%IHU, Strasbourg, France%
%}
%
\maketitle

@@

\keywords{liver, Glisson's capsule, vascular network, deformation modeling, membrane}
\end{abstract}

@@

It is known that the parenchyma exhibits viscoelastic behaviour \TG{Cite
something}, we however employ simpler elastic model as we are not so much
interested in time-dependent behaviour but rather in static equilibrium
under certain conditions.
We also rely on the vascularized model of the parenchyma proposed by
Peterl\'{i}k et al. \cite{Peterlik2012}.

The parenchyma is modeled using corotational finite elements. The corotational
formulation allows us to deal with geometrical non-linearities but still
rely on linear stress-strain relationship. Thus rotations and large
deformations are possible. In the corotational formulation the deformation
of each element is expressed in the local frame of reference. There are
several ways of computing the corotational frame for elements, we rely on
the geometrical method proposed in \cite{Nesme2005}.
% NOTE: This description of corotational method is very simplified and could be extended.
The stiffness matrix $\Mat{K}$ can be expressed in terms of the deformation
$u$ and the force as $\Vec{f} = \Mat{K}(\Vec{u}) \Vec{u}$. Where the
components of $\Mat{K}$ depend on the orientation of each element in the
current simulation step.

The vascularization method is based on linear beams
with local frames of reference (in a many aspects similar to the
corotational formulation). As such the model also handles geometric
non-linearities in the deformation. The proposed model assumes no relative
motion of vessels to the parenchyma. Segmented vascular network is
discretized into the set of linear elements that are then projected into
the tetrahedral mesh of the parenchyma. During the simulation forces and
torques resulting from the deformations of vessels by parenchyma are
computed and propagated back onto the tetrahedral mesh.

@@

@@

\CD{If the experiments demonstrate that the stiffness of the material is different, the thickness is very very small... so we have to evaluate if it is really necessary to include the model of the capsule for both local and global deformations.}
\TG{I didn't undersand that. What do you mean?}

\subsection{Local Deformations} %{{{

During the contact with an instrument (probe, needle insertion, sharp
instruments, \ldots) a specific deformation occurs in the near vicinity of
the instrument. This type of deformation may not necessarily induce the
deformation of the object as a whole and therefore can be considered as
local. Correct material properties are not only important for amount of the
displacement but play also a crucial role in capturing the correct area of
the deformation or its profile near the instrument.
\TG{Does it make sense this way? I can add an image. :)}

One example of such local deformation occurs during the aspiration test
when the response of liver to the negative pressure is measured.
Hollenstein et al. \cite{Hollenstein2006} observed that the Glisson's
capsule has non-negligible influence on response of the tissue and modeling
only the parenchyma leads to overestimation of the deformation. We
reproduced the test with our model and acquired similar results.

The aspiration device consists of a tube with 1 cm in diameter and is able
to control internal pressure in the tube. The test is performed by
attaching the tube to the tissue and measuring the tissue response. We
have set up a simulation in the spirit of this test. 
The simulation is performed with a 15x15x15 mm cube of liver tissue and tube
with 1 cm in diameter. The pressure of 3 kPa is simulated inside the tube.

Figure \ref{fig-aspiration} shows the profiles of cuts in the middle of the
test cube. It can be seen that the
deformation for the model without capsule is much larger. This is in good
agreement with the data obtained by Hollenstein et al.
